---
title: "Auditory_FB_analyses"
author: "Lucrezia Lonardo"
date: "2024-10-22"
output: html_document
R_version: 4.2.2
---
### Load libraries and custom functions
```{r}
rm(list=ls())
library(tidyverse)
library(ggplot2)
library(lme4)
library(summarytools)
library(car)
source("functions/diagnostic_fcns.r")
source("functions/glmm_stability.r")
source("functions/boot_glmm.r")
```

### Import data

```{r}
xdata <- read.csv("data/tested_dogs.csv", header = TRUE) %>% 
   mutate(choice_letter = as.factor(choice_letter),
         condition = as.factor(condition),
         age_months = as.numeric(age_months)) %>% 
  filter(choice_letter %in% c("A", "B")) %>%   # Keep rows where choice_letter is "A" or "B" in case of NAs
  droplevels()

which(is.na(xdata$choice_letter)) #double check for NAs
```

###Descriptive statistics
```{r}
##nr. dogs choosing bucket A and B overall
table(xdata$choice_letter)

##proportions relative to the whole sample
mean(xdata$choice_letter=="A")
mean(xdata$choice_letter=="B")

##nr. dogs choosing bucket A and B in each condition
table(xdata$condition, xdata$choice_letter)

##proportions choices relative to the different conditions
fb_subset <- xdata[xdata$condition == "false_belief", ]
know_subset <- xdata[xdata$condition == "knowledge", ]
ctr_subset <- xdata[xdata$condition == "control", ]
mean(fb_subset$choice_letter=="A")
mean(know_subset$choice_letter=="A")
mean(ctr_subset$choice_letter=="A")

#experimenters (communicator and hider)
table(xdata$condition, xdata$communicator)
table(xdata$condition, xdata$hider)
xdata$experimenter_dyad<-paste(xdata$communicator, xdata$hider, sep="_")
table(xdata$condition, xdata$experimenter_dyad)
#choice and experimenters
table(xdata$choice_letter, xdata$experimenter_dyad)

#age
mean(fb_subset$age_months)
mean(know_subset$age_months) 
mean(ctr_subset$age_months)
#how many <2 years in each condition
sum(fb_subset$age_months<24) #9
sum(know_subset$age_months<24) #9
sum(ctr_subset$age_months<24) #5


#sex
table(xdata$sex, xdata$condition)
#FCI group
table(xdata$FCI_group, xdata$condition)
table(xdata$condition)
#choice and FCI group
table(xdata$choice_letter, xdata$FCI_group)
```


### Plots

```{r pressure, echo=FALSE}
#table with descriptive stats for plotting
plot.data <- xdata %>%
  mutate(choice = ifelse(choice_letter == "A", 1, 0)) %>%
  mutate(condition = fct_relevel(
    fct_recode(
      condition,
      "false belief" = "false_belief",
      "knowledge" = "knowledge",
      "control" = "control"
    ),
    "false belief",
    "knowledge",
    "control"
  )) %>% #specify order of factor levels for factor_relevel
  group_by(condition) %>%
  summarise(
    mean.choiceA = mean(choice),
    sd.choiceA = sd(choice),
    n.choseA = sum(choice),
    median.choiceA = median(choice),
    n.tested.per.cond = length(choice),
    se.choiceA = sd(choice) / sqrt(length(choice))
  )

plot.data

##barplot of choices by experimental group

p.choice <- ggplot(data = plot.data, aes(x = condition, y = mean.choiceA)) +
  geom_bar(stat = "identity", alpha = 0.5) +
  geom_errorbar(aes(ymin = mean.choiceA - se.choiceA, ymax = mean.choiceA + #error bars represent standard error
                      se.choiceA),
                width = .2) +
  #geom_signif(comparisons=list(c("False Belief", #"Knowledge")), annotations="*",
  #             y_position = 0.75, tip_length = 0.04, #vjust=0.4) +
  #ylim(0.2,0.6)+
  theme_classic() +
  labs(x = "Experimental group", y = "Proportion of bucket A choices")+
  ggtitle("All breeds (N=148)")

p.choice
 
#saving the plot with ggsave
# ggsave(
#   "./graphs/bucket_A_choices_BC.png",
#   plot = last_plot(),
#   device = NULL,
#  # path = "C:/Users/lonardol/Desktop/False Belief experiment/FB_2_follow_up/Graphs/Border Collies",
#   scale = 0.8,
#   width = 7,
#   height = 6,
#   #units = c("in", "cm", "mm"),
#   dpi = 320,
#   limitsize = TRUE
# )
```

```{r}
#mosaic plot to visualize also bucket B choices
mosaic.plot.data  <- xdata %>%
  mutate(choice_letter = fct_relevel(choice_letter, "B", "A")) %>%
  mutate(condition = fct_relevel(
    fct_recode(
      condition,
      "False belief" = "false_belief",
      "True Belief" = "knowledge",
      "Sound Control" = "control"
    ),
    "False belief",
    "True Belief",
    "Sound Control"
  )) #specify order of factor levels for factor_relevel

plot(mosaic.plot.data$condition,mosaic.plot.data$choice_letter,
  xlab = "Condition",
  ylab = "Bucket chosen")


#saving the plot with ggsave
# ggsave(
#   "./graphs/bucket_A_choices_BC.png",
#   plot = last_plot(),
#   device = NULL,
#  # path = "C:/Users/lonardol/Desktop/False Belief experiment/FB_2_follow_up/Graphs/Border Collies",
#   scale = 0.8,
#   width = 7,
#   height = 6,
#   #units = c("in", "cm", "mm"),
#   dpi = 320,
#   limitsize = TRUE
# )
```
```{r}
#chi square test (false belief - knowledge) 
condition<-c("false belief", "knowledge", "control")
outcome <- plot.data$n.choseA # number of successes (bucket A chosen) for each condition
n <- plot.data$n.tested.per.cond # total trials for each condition

# Create a contingency table
table <- matrix(c(outcome[1], n[1] - outcome[1], outcome[2], n[2] - outcome[2], outcome[3], n[3] - outcome[3]), 
                nrow = 3, byrow = TRUE)
rownames(table) <- condition
colnames(table) <- c("Success", "Failure")

# Chi-squared test
chisq.test(table)
```
```{r}
#plot with error bars showing confidence intervals 
plot.data.ci <- plot.data %>%
  mutate(ci.lower = mean.choiceA - 1.96 * se.choiceA,
         ci.upper = mean.choiceA + 1.96 * se.choiceA)

# Now, plot with confidence intervals as error bars
p.choice.ci <- ggplot(data = plot.data.ci, aes(x = condition, y = mean.choiceA)) +
  geom_bar(stat = "identity", alpha = 0.5) +
  geom_errorbar(aes(ymin = ci.lower, ymax = ci.upper), width = 0.2) +
  #ylim(0,0.8)+
  #geom_signif(comparisons=list(c("False Belief", "Knowledge")), annotations="*",
  #             y_position = 0.75, tip_length = 0.04, vjust=0.4) +
  theme_classic() +
  labs(x = "Experimental group", y = "Proportion of bucket A choices") +
  ggtitle("All breeds (N=148)")

p.choice.ci
```
### Preparing data for modelling
```{r}
model.data<-xdata %>% 
  #making sure factors are treated as such
  mutate(dog_ID=as.factor(dog_ID)) %>% 
  mutate(breed=as.factor(breed)) %>% 
  mutate(sex=as.factor(sex)) %>% 
  mutate(first_baited_box=as.factor(first_bucket_A)) %>% 
  select(-first_bucket_A) %>% 
  mutate(FCI_group=as.factor(FCI_group)) %>% 
  mutate(experience_in_dog_sport=as.factor(experience_in_dog_sport)) %>%
  mutate(hider=as.factor(hider)) %>%
  mutate(communicator=as.factor(communicator)) %>%
  mutate(experimenter_dyad=as.factor(experimenter_dyad)) %>%
  mutate(choice_colour=as.factor(choice_colour)) %>% 
  #centering variables
  mutate(z.age = as.numeric(scale(age_months, scale = T, center = T)), #age scaled and centered (=z-transformed, we subtracted the mean and divided by SD)
         sex.c = as.numeric(scale(as.numeric(as.factor(sex)), scale=F, center=T)), #factors manually dummy coded and centered (i.e., we subtracted the mean value from each observed value)
         condition.c = as.numeric(scale(as.numeric(as.factor(condition)), scale=F, center=T)),
         first_baited_box.c = as.numeric(scale(as.numeric(as.factor(first_baited_box)), scale=F, center=T)))



view(dfSummary(model.data))

summary(model.data$age_months) #23rd December 2024: mean age: 61 months (ca. 5 years); min age: 6 months; max age: 155 months (12.9 years)
#which(model.data$age_months>156) #check in case the distribution has unlikely outliers.

#checking familiarisation performance is as expected (max 4 trials per phase)
which(model.data$nr_fam_phase1_trials==5) #there is one exception: Garou because one trial in phase 1 had to be repeated due to owner's mistake

#checking no choice trials
which(model.data$no_choice_trials=='')
#Heidi6's familiarisation performance was not reported in the data file. Please correct.
which(model.data$no_choice_trials=="1_in_testtrial")

#checking which breeds we have
unique(levels(model.data$breed)) #some levels are redundant (e.g. schanuzer and giant schnauzer; belgian shepherd and malinois), so let's reduce unnecessary variation between breeds


model.data <- model.data %>% 
  mutate(breed_recoded = case_when(
    breed == "malinois" | breed == "laekenois" ~ "belgian_shepherd",
    breed == "giant_schnauzer" ~ "schnauzer",
    grepl("corgi", breed) ~ "welsh_corgi",
    TRUE ~ as.character(breed) # Preserve original values as character
  ),
  breed_recoded = as.factor(breed_recoded))

#check new levels of breed 
unique(levels(model.data$breed_recoded)) #45 different breeds
#what is a miniature american shepherd? Not in FCI!

#checking if the data preparations worked as expected
str(model.data)
```
### GLMM 01: Model with breed as random intercept 

```{r eval=FALSE, echo=FALSE}
xx.fe.re=fe.re.tab(fe.model="choice_letter ~ condition + age_months + sex + first_baited_box",
                   re="(1|breed_recoded)", data=model.data)
xx.fe.re$summary

#exclude all rs within breed at least for now until we have all data?
```

``` {r mixed modeling, error=TRUE}
#full model with all rs within breed. Right now (23 December 2024), we don't have enough variation within breeds for all these parameters, so we remove the parameters for the correlations between random slopes and intercept. 

# full.choice.breed <- glmer(choice_letter ~ condition + z.age + sex + first_baited_box+
#              (1+ sex.c + condition.c + first_baited_box.c + z.age | breed.recoded), 
#              data = model.data, family = binomial, 
#              control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5))
# )

full.choice.breed <- glmer(choice_letter ~ condition + z.age + sex + first_baited_box+
             (1 + condition.c +  z.age + sex.c + first_baited_box.c || breed_recoded), 
             data = model.data, family = binomial, 
             control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5))
)
```

#### Distribution of BLUPs
```{r}
ranef.diagn.plot(full.choice.breed)
```

#### model output
  + Coefficients
```{r}
round(summary(full.choice.breed)$coefficients, 3)
```
  + Individual predictors: likelihood ratio tests  
Drop1: P values for the individual effects were based on likelihood ratio tests comparing the full with respective reduced models (Barr et al., 2013; R function drop1 with argument 'test' set to "Chisq"). 

```{r}
drop1.full=drop1(full.choice.breed, test="Chisq")
round(drop1.full,3)
```
```{r}
summary(full.choice.breed)$varcor
```
#### confidence intervals

```{r}
boot.res=boot.glmm.pred(model.res=full.choice.breed, excl.warnings=T,
	nboots=1000, para=T)
res<-round(boot.res$ci.estimates, 3)
res
```
#### model stability

```{r eval=FALSE, include=FALSE}

# One breed at a time excluded to assess the impact of possible outliers 
m.stab.b <- glmm.model.stab(model.res = full.choice.breed, use = c("breed_recoded"))

table(m.stab.b$detailed$lme4.warnings) #some models failed to converge

xx <- as.data.frame(round(m.stab.b$summary[, -1], 3))
#table with model stability
xx

#plot of stability (remember to execute all three lines at the same time)
png("plots/bin_glmm01_choice_stability.png")
m.stab.plot(round(m.stab.b$summary[, -1], 3))
dev.off()
```