---
title: "Auditory_FB_analyses"
author: "Lucrezia Lonardo"
date: "2024-10-22"
output: html_document
R_version: 4.2.2
---
### Load libraries and custom functions
```{r}
rm(list=ls())
library(tidyverse)
library(ggplot2)
library(lme4)
library(summarytools)
library(car)
library(writexl)
library(readxl)
library(irr)
source("functions/diagnostic_fcns.r")
source("functions/glmm_stability.r")
source("functions/boot_glmm.r")
```

### Import data

```{r}
xdata <- read.csv("data/tested_dogs.csv", header = TRUE) %>% 
   mutate(choice_letter = as.factor(choice_letter),
         condition = as.factor(condition),
         age_months = as.numeric(age_months)) %>% 
  filter(choice_letter %in% c("A", "B")) %>%   # Keep rows where choice_letter is "A" or "B" in case of NAs
  droplevels()

which(is.na(xdata$choice_letter)) #double check for NAs
```

###Descriptive statistics
```{r}
table(xdata$condition)
##nr. dogs choosing bucket A and B overall
table(xdata$choice_letter)

##proportions relative to the whole sample
mean(xdata$choice_letter=="A")
mean(xdata$choice_letter=="B")

#binomial test to check if proportion of choice A and B differ significantly overall across conditions
binom.test(101, 240, p=0.5)

##nr. dogs choosing bucket A and B in each condition
table(xdata$condition, xdata$choice_letter)

##proportions choices relative to the different conditions
fb_subset <- xdata[xdata$condition == "false_belief",]
know_subset <- xdata[xdata$condition == "knowledge", ]
ctr_subset <- xdata[xdata$condition == "control", ]
mean(fb_subset$choice_letter=="A")
mean(know_subset$choice_letter=="A")
mean(ctr_subset$choice_letter=="A")

#experimenters (communicator and hider)
table(xdata$condition, xdata$communicator)
table(xdata$condition, xdata$hider)
xdata$experimenter_dyad<-paste(xdata$communicator, xdata$hider, sep="_")
table(xdata$condition, xdata$experimenter_dyad)
#choice and experimenters
table(xdata$choice_letter, xdata$experimenter_dyad)

#age
range(xdata$age)
mean(xdata$age)
mean(fb_subset$age_months)
mean(know_subset$age_months, na.rm=T) 
mean(ctr_subset$age_months)
#how many <2 years in each condition
sum(fb_subset$age_months<24) #9
sum(know_subset$age_months<24, na.rm=T) #15
sum(ctr_subset$age_months<24) #7


#sex
table(xdata$sex)
table(xdata$sex, xdata$condition)
#FCI group
table(xdata$FCI_group, xdata$condition)
#choice and FCI group
table(xdata$choice_letter, xdata$FCI_group)

#breed recoded
xdata <- xdata %>% 
  mutate(breed_recoded = case_when(
    breed == "malinois" | breed == "laekenois" ~ "belgian_shepherd",
    breed == "giant_schnauzer" ~ "schnauzer",
    breed == "deutsch_langhaar" ~ "german_longhaired_pointer",
    breed == "samojede" ~ "samoyed",
    grepl("corgi", breed) ~ "welsh_corgi",
    TRUE ~ as.character(breed) # Preserve original values as character
  ),
  breed_recoded = as.factor(breed_recoded))

#check new levels of breed 
unique(levels(xdata$breed_recoded)) #61 breeds

#breed
table(xdata$condition, xdata$breed_recoded)
#table(xdata$breed_recoded, xdata$sex)
unique(levels(xdata$breed_recoded))

#food
sum(grepl(c("sausages"), xdata$notes))
sum(grepl(c("own food"), xdata$notes))
```


### Plots

```{r pressure, echo=FALSE}
#table with descriptive stats for plotting
plot.data <- xdata %>%
  mutate(choice = ifelse(choice_letter == "A", 1, 0)) %>%
  mutate(condition = fct_relevel(
    fct_recode(
      condition,
      "false belief" = "false_belief",
      "knowledge" = "knowledge",
      "control" = "control"
    ),
    "false belief",
    "knowledge",
    "control"
  )) %>% #specify order of factor levels for factor_relevel
  group_by(condition) %>% 
  summarise(
    mean.choiceA = mean(choice),
    sd.choiceA = sd(choice),
    n.choseA = sum(choice),
    median.choiceA = median(choice),
    n.tested.per.cond = length(choice),
    se.choiceA = sd(choice) / sqrt(length(choice))
  )

##barplot of choices by experimental group

p.choice <- ggplot(data = plot.data, aes(x = condition, y = mean.choiceA)) +
  geom_bar(stat = "identity", alpha = 0.5) +
  geom_errorbar(aes(ymin = mean.choiceA - se.choiceA, ymax = mean.choiceA + #error bars represent standard error
                      se.choiceA),
                width = .2) +
  #geom_signif(comparisons=list(c("False Belief", #"Knowledge")), annotations="*",
  #             y_position = 0.75, tip_length = 0.04, #vjust=0.4) +
  #ylim(0.2,0.6)+
  theme_classic() +
  labs(x = "Experimental group", y = "Proportion of bucket A choices")+
  ggtitle("All breeds (N=240)")

p.choice
 
#saving the plot with ggsave
# ggsave(
#   "./graphs/bucket_A_choices_BC.png",
#   plot = last_plot(),
#   device = NULL,
#  # path = "C:/Users/lonardol/Desktop/False Belief experiment/FB_2_follow_up/Graphs/Border Collies",
#   scale = 0.8,
#   width = 7,
#   height = 6,
#   #units = c("in", "cm", "mm"),
#   dpi = 320,
#   limitsize = TRUE
# )
```

```{r}
#mosaic plot to visualize also bucket B choices
mosaic.plot.data  <- xdata %>%
  mutate(choice_letter = fct_relevel(choice_letter, "B", "A")) %>%
  mutate(condition = fct_relevel(
    fct_recode(
      condition,
      "False belief" = "false_belief",
      "True Belief" = "knowledge",
      "Sound Control" = "control"
    ),
    "False belief",
    "True Belief",
    "Sound Control"
  )) #specify order of factor levels for factor_relevel

png("./plots/mosaic_choices_240_dogs.png")
plot(mosaic.plot.data$condition,mosaic.plot.data$choice_letter,
  xlab = "Condition",
  ylab = "Bucket chosen")
dev.off()

#saving the plot with ggsave
# ggsave(
#   "./plots/mosaic_choices_168_dogs.png",
#   plot = last_plot(),
#   device = NULL,
#  # path = "C:/Users/lonardol/Desktop/False Belief experiment/FB_2_follow_up/Graphs/Border Collies",
#   scale = 0.8,
#   width = 7,
#   height = 6,
#   #units = c("in", "cm", "mm"),
#   dpi = 320,
#   limitsize = TRUE
# )
```
```{r}
#add column with independent vs cooperative breeds distinction 
xdata<-xdata %>% 
  mutate(breed_type = ifelse(FCI_group %in% c("1", "2", "7", "8"), "cooperative", "independent"))

table(xdata$breed_type)

table(xdata$choice_letter, xdata$breed_type, xdata$condition)

#add column with experience in dogs sport as binary (yes/no) variable
xdata<-xdata %>% 
  mutate(training = ifelse(experience_in_dog_sport=="no", "no", "yes"))

table(xdata$training)

table(xdata$FCI_group[xdata$training=="no"])

table(xdata$choice_letter, xdata$training, xdata$condition)

#experimenter ID
table(xdata$communicator, xdata$choice_letter, xdata$condition)
```
### Plot breed performance
```{r}
#breed_type data
cooperative_data<-xdata %>% 
  filter(breed_type!="independent")
independent_data<-xdata %>% 
  filter(breed_type!="cooperative")

#mosaic plot for cooperative
mosaic.cooperative.plot.data  <- cooperative_data %>%
  mutate(choice_letter = fct_relevel(choice_letter, "B", "A")) %>%
  mutate(condition = fct_relevel(
    fct_recode(
      condition,
      "False belief" = "false_belief",
      "True Belief" = "knowledge",
      "Sound Control" = "control"
    ),
    "False belief",
    "True Belief",
    "Sound Control"
  )) #specify order of factor levels for factor_relevel

png("./plots/mosaic_choices_206_dogs_cooperative.png")
plot(mosaic.cooperative.plot.data$condition,mosaic.cooperative.plot.data$choice_letter,
  xlab = "Condition",
  ylab = "Bucket chosen",
  main = "Cooperative breeds" )
dev.off()

#mosaic plot for independent
mosaic.independent.plot.data  <- independent_data %>%
  mutate(choice_letter = fct_relevel(choice_letter, "B", "A")) %>%
  mutate(condition = fct_relevel(
    fct_recode(
      condition,
      "False belief" = "false_belief",
      "True Belief" = "knowledge",
      "Sound Control" = "control"
    ),
    "False belief",
    "True Belief",
    "Sound Control"
  )) #specify order of factor levels for factor_relevel

png("./plots/mosaic_choices_206_dogs_independent.png")
plot(mosaic.independent.plot.data$condition,mosaic.independent.plot.data$choice_letter,
  xlab = "Condition",
  ylab = "Bucket chosen",
  main = "Independent breeds" )
dev.off()

#subset the data to the different FCI groups
FCI1_data<-xdata %>% 
  filter(FCI_group=="1")

FCI2_data<-xdata %>% 
  filter(FCI_group=="2")

FCI5_data<-xdata %>% 
  filter(FCI_group=="5")

FCI6_data<-xdata %>% 
  filter(FCI_group=="6")

FCI7_data<-xdata %>% 
  filter(FCI_group=="7")

FCI8_data<-xdata %>% 
  filter(FCI_group=="8")

FCI10_data<-xdata %>% 
  filter(FCI_group=="10")

# mosaic plot for FCI group 1
mosaic.FCI1.plot.data  <- FCI1_data %>%
  mutate(choice_letter = fct_relevel(choice_letter, "B", "A")) %>%
  mutate(condition = fct_relevel(
    fct_recode(
      condition,
      "False belief" = "false_belief",
      "True Belief" = "knowledge",
      "Sound Control" = "control"
    ),
    "False belief",
    "True Belief",
    "Sound Control"
  )) #specify order of factor levels for factor_relevel

png("./plots/FCI_groups/mosaic_choices_206_dogs_FCI1.png")
plot(mosaic.FCI1.plot.data$condition, mosaic.FCI1.plot.data$choice_letter,
  xlab = "Condition",
  ylab = "Bucket chosen",
  main = "FCI group 1 (N=84)" )
dev.off()

# mosaic plot for FCI group 2
mosaic.FCI2.plot.data  <- FCI2_data %>%
  mutate(choice_letter = fct_relevel(choice_letter, "B", "A")) %>%
  mutate(condition = fct_relevel(
    fct_recode(
      condition,
      "False belief" = "false_belief",
      "True Belief" = "knowledge",
      "Sound Control" = "control"
    ),
    "False belief",
    "True Belief",
    "Sound Control"
  )) #specify order of factor levels for factor_relevel

png("./plots/FCI_groups/mosaic_choices_206_dogs_FCI2.png")
plot(mosaic.FCI5.plot.data$condition, mosaic.FCI5.plot.data$choice_letter,
  xlab = "Condition",
  ylab = "Bucket chosen",
  main = "FCI group 2 (N=11)" )
dev.off()

# mosaic plot for FCI group 5
mosaic.FCI5.plot.data  <- FCI5_data %>%
  mutate(choice_letter = fct_relevel(choice_letter, "B", "A")) %>%
  mutate(condition = fct_relevel(
    fct_recode(
      condition,
      "False belief" = "false_belief",
      "True Belief" = "knowledge",
      "Sound Control" = "control"
    ),
    "False belief",
    "True Belief",
    "Sound Control"
  )) #specify order of factor levels for factor_relevel

png("./plots/FCI_groups/mosaic_choices_206_dogs_FCI5.png")
plot(mosaic.FCI5.plot.data$condition, mosaic.FCI5.plot.data$choice_letter,
  xlab = "Condition",
  ylab = "Bucket chosen",
  main = "FCI group 5 (N=11)" )
dev.off()

# mosaic plot for FCI group 6
mosaic.FCI6.plot.data  <- FCI6_data %>%
  mutate(choice_letter = fct_relevel(choice_letter, "B", "A")) %>%
  mutate(condition = fct_relevel(
    fct_recode(
      condition,
      "False belief" = "false_belief",
      "True Belief" = "knowledge",
      "Sound Control" = "control"
    ),
    "False belief",
    "True Belief",
    "Sound Control"
  )) #specify order of factor levels for factor_relevel

png("./plots/FCI_groups/mosaic_choices_206_dogs_FCI6.png")
plot(mosaic.FCI6.plot.data$condition, mosaic.FCI6.plot.data$choice_letter,
  xlab = "Condition",
  ylab = "Bucket chosen",
  main = "FCI group 6 (N=6)" )
dev.off()

# mosaic plot for FCI group 7
mosaic.FCI7.plot.data  <- FCI7_data %>%
  mutate(choice_letter = fct_relevel(choice_letter, "B", "A")) %>%
  mutate(condition = fct_relevel(
    fct_recode(
      condition,
      "False belief" = "false_belief",
      "True Belief" = "knowledge",
      "Sound Control" = "control"
    ),
    "False belief",
    "True Belief",
    "Sound Control"
  )) #specify order of factor levels for factor_relevel

png("./plots/FCI_groups/mosaic_choices_206_dogs_FCI7.png")
plot(mosaic.FCI7.plot.data$condition, mosaic.FCI7.plot.data$choice_letter,
  xlab = "Condition",
  ylab = "Bucket chosen",
  main = "FCI group 7 (N=13)" )
dev.off()

# mosaic plot for FCI group 8
mosaic.FCI8.plot.data  <- FCI8_data %>%
  mutate(choice_letter = fct_relevel(choice_letter, "B", "A")) %>%
  mutate(condition = fct_relevel(
    fct_recode(
      condition,
      "False belief" = "false_belief",
      "True Belief" = "knowledge",
      "Sound Control" = "control"
    ),
    "False belief",
    "True Belief",
    "Sound Control"
  )) #specify order of factor levels for factor_relevel

png("./plots/FCI_groups/mosaic_choices_206_dogs_FCI8.png")
plot(mosaic.FCI8.plot.data$condition, mosaic.FCI8.plot.data$choice_letter,
  xlab = "Condition",
  ylab = "Bucket chosen",
  main = "FCI group 8 (N=64)" )
dev.off()

# mosaic plot for FCI group 10
mosaic.FCI10.plot.data  <- FCI10_data %>%
  mutate(choice_letter = fct_relevel(choice_letter, "B", "A")) %>%
  mutate(condition = fct_relevel(
    fct_recode(
      condition,
      "False belief" = "false_belief",
      "True Belief" = "knowledge",
      "Sound Control" = "control"
    ),
    "False belief",
    "True Belief",
    "Sound Control"
  )) #specify order of factor levels for factor_relevel

png("./plots/FCI_groups/mosaic_choices_206_dogs_FCI10.png")
plot(mosaic.FCI10.plot.data$condition, mosaic.FCI10.plot.data$choice_letter,
  xlab = "Condition",
  ylab = "Bucket chosen",
  main = "FCI group 10 (N=6)" )
dev.off()
```

```{r}
#plot with error bars showing confidence intervals 
plot.data.ci <- plot.data %>%
  mutate(ci.lower = mean.choiceA - 1.96 * se.choiceA,
         ci.upper = mean.choiceA + 1.96 * se.choiceA)

# Now, plot with confidence intervals as error bars
p.choice.ci <- ggplot(data = plot.data.ci, aes(x = condition, y = mean.choiceA)) +
  geom_bar(stat = "identity", alpha = 0.5) +
  geom_errorbar(aes(ymin = ci.lower, ymax = ci.upper), width = 0.2) +
  #ylim(0,0.8)+
  #geom_signif(comparisons=list(c("False Belief", "Knowledge")), annotations="*",
  #             y_position = 0.75, tip_length = 0.04, vjust=0.4) +
  theme_classic() +
  labs(x = "Experimental group", y = "Proportion of bucket A choices") +
  ggtitle("All breeds (N=240)")

p.choice.ci
```

### Preparing data for modelling
```{r}
model.data<-xdata %>% 
  #making sure factors are treated as such
  mutate(dog_ID=as.factor(dog_ID)) %>% 
  mutate(breed=as.factor(breed)) %>% 
  mutate(breed_recoded=as.factor(breed_recoded)) %>% 
  mutate(breed_type=as.factor(breed_type)) %>% 
  mutate(sex=as.factor(sex)) %>% 
  mutate(first_baited_box=as.factor(first_bucket_A)) %>% 
  select(-first_bucket_A) %>% 
  mutate(FCI_group=as.factor(FCI_group)) %>% 
  mutate(training=as.factor(training)) %>%
  mutate(hider=as.factor(hider)) %>%
  mutate(communicator=as.factor(communicator)) %>%
  mutate(experimenter_dyad=as.factor(experimenter_dyad)) %>%
  mutate(choice_colour=as.factor(choice_colour)) %>% 
  #centering variables
  mutate(z.age = as.numeric(scale(age_months, scale = T, center = T)), #age scaled and centered (=z-transformed, we subtracted the mean and divided by SD)
         sex.c = as.numeric(scale(as.numeric(as.factor(sex)), scale=F, center=T)), #factors manually dummy coded and centered (i.e., we subtracted the mean value from each observed value)
         condition.c = as.numeric(scale(as.numeric(as.factor(condition)), scale=F, center=T)),
         breed_type.c = as.numeric(scale(as.numeric(as.factor(breed_type)), scale=F, center=T)),
         first_baited_box.c = as.numeric(scale(as.numeric(as.factor(first_baited_box)), scale=F, center=T)),
         training.c = as.numeric(scale(as.numeric(as.factor(training)), scale=F, center=T)))



view(dfSummary(model.data))

summary(model.data$age_months)

#checking familiarisation performance is as expected (max 4 trials per phase)
which(model.data$nr_fam_phase1_trials>4) #there is one exception: Garou because one trial in phase 1 had to be repeated due to owner's mistake

#checking no choice trials
which(model.data$no_choice_trials=='')
#Heidi6's familiarisation performance was not reported in the data file. Please correct.
which(model.data$no_choice_trials=="1_in_testtrial")

#check new levels of breed 
unique(levels(model.data$breed_recoded)) #61 different breeds

#checking if the data preparations worked as expected
str(model.data)
```
### GLMM 01: Model with breed as random intercept 

```{r eval=FALSE, echo=FALSE}
xx.fe.re=fe.re.tab(fe.model="choice_letter ~ condition + age_months + sex + first_baited_box",
                   re="(1|breed_recoded)", data=model.data)
xx.fe.re$summary
```

``` {r mixed modeling, error=TRUE}
#full model with all rs within breed. We don't have enough variation within breeds for all these parameters, so we remove the parameters for the correlations between random slopes and intercept. 

full.choice <- glmer(choice_letter ~ condition + z.age + sex + first_baited_box+
             (1 + condition.c +  z.age + sex.c + first_baited_box.c || breed_recoded), 
             data = model.data, family = binomial, 
             control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5))
)
```
#### Variance Inflation Factors
```{r}
vif(full.choice) #collinearity is no problem
```

#### Distribution of BLUPs
```{r}
ranef.diagn.plot(full.choice)
```

#### model output
  + Coefficients
```{r}
round(summary(full.choice)$coefficients, 3)
```
  + Individual predictors: likelihood ratio tests  
Drop1: P values for the individual effects were based on likelihood ratio tests comparing the full with respective reduced models (Barr et al., 2013; R function drop1 with argument 'test' set to "Chisq"). 

```{r}
drop1.full=drop1(full.choice, test="Chisq")
round(drop1.full,3)
```
```{r}
summary(full.choice)$varcor
```
#### confidence intervals

```{r}
boot.res=boot.glmm.pred(model.res=full.choice, excl.warnings=T,
	nboots=1000, para=T)
res<-round(boot.res$ci.estimates, 3)
res
```
#### model stability

```{r eval=FALSE, include=FALSE}

# One breed at a time excluded to assess the impact of possible outliers 
m.stab.b <- glmm.model.stab(model.res = full.choice, use = c("breed_recoded"))

table(m.stab.b$detailed$lme4.warnings) #some models failed to converge

xx <- as.data.frame(round(m.stab.b$summary[, -1], 3))
#table with model stability
xx

#plot of stability (execute all three lines at the same time)
png("plots/bin_glmm01_choice_stability.png")
m.stab.plot(round(m.stab.b$summary[, -1], 3))
dev.off()
```


### GLMM 02: Model with interaction breed_type(coop. vs indep.) * condition (Exploratory)

```{r eval=FALSE, echo=FALSE}
xx.fe.re=fe.re.tab(fe.model="choice_letter ~ breed_type*condition + age_months + sex + first_baited_box",
                   re="(1|breed_recoded)", data=model.data)
xx.fe.re$summary
```

``` {r mixed modeling, error=TRUE}
#full model with all rs within breed. Right now (23 December 2024), we don't have enough variation within breeds for all these parameters, so we remove the parameters for the correlations between random slopes and intercept. 

full.choice.int <- glmer(choice_letter ~ condition*breed_type + z.age + sex + first_baited_box+
             (1+ sex.c + condition.c*breed_type.c + first_baited_box.c + z.age || breed_recoded),
             data = model.data, family = binomial,
             control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5))
)

#had to exclude correlation parameters between rss and icpt because they were unidentifiable
```


#### model output
  + Coefficients
```{r}
round(summary(full.choice.int)$coefficients, 3)
```
  + Individual predictors: likelihood ratio tests  
Drop1: P values for the individual effects were based on likelihood ratio tests comparing the full with respective reduced models (Barr et al., 2013; R function drop1 with argument 'test' set to "Chisq"). 

```{r}
drop1.full=drop1(full.choice.int, test="Chisq")
round(drop1.full,3)
#interaction not significant
```
####the interaction condition*breed_type is not significant, so we remove it but keep both main effects
``` {r mixed modeling, error=TRUE}
#full model with all rs within breed. Right now (23 December 2024), we don't have enough variation within breeds for all these parameters, so we remove the parameters for the correlations between random slopes and intercept. 

full.choice.breed <- glmer(choice_letter ~ condition + breed_type + z.age + sex + first_baited_box+
             (1+ sex.c + condition.c + breed_type.c + first_baited_box.c + z.age || breed_recoded),
             data = model.data, family = binomial,
             control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5))
)

#had to exclude correlation parameters between rss and icpt because they were unidentifiable
```
#### Variance Inflation Factors
```{r}
vif(full.choice.breed) #collinearity is no problem
```

#### Distribution of BLUPs
```{r}
ranef.diagn.plot(full.choice.breed)
```
#### model output
  + Coefficients
```{r}
round(summary(full.choice.breed)$coefficients, 3)
```
  + Individual predictors: likelihood ratio tests  
Drop1: P values for the individual effects were based on likelihood ratio tests comparing the full with respective reduced models (Barr et al., 2013; R function drop1 with argument 'test' set to "Chisq"). 

```{r}
drop1.full=drop1(full.choice.breed, test="Chisq")
round(drop1.full,3)
#the interaction condition*breed_type is not significant, so we remove it but keep both main effects
```
#### confidence intervals

```{r}
boot.res.breed=boot.glmm.pred(model.res=full.choice.breed, excl.warnings=T,
	nboots=1000, para=T)
res.breed<-round(boot.res.breed$ci.estimates, 3)
res.breed
```
#### model stability

```{r eval=FALSE, include=FALSE}

# One breed at a time excluded to assess the impact of possible outliers 
m.stab.breed <- glmm.model.stab(model.res = full.choice.breed, use = c("breed_recoded"))

table(m.stab.breed$detailed$lme4.warnings) #several models failed to converge

xx.breed <- as.data.frame(round(m.stab.b$summary[, -1], 3))
#table with model stability
xx.breed

#plot of stability (remember to execute all three lines at the same time)
png("plots/bin_glmm02_breed_type_stability.png")
m.stab.plot(round(m.stab.breed$summary[, -1], 3))
dev.off()
```
### GLMM 03: Effect of training*condition

```{r eval=FALSE, echo=FALSE}
xx.fe.re=fe.re.tab(fe.model="choice_letter ~ breed_type*condition + age_months + sex + first_baited_box",
                   re="(1|breed_recoded)", data=model.data)
xx.fe.re$summary
```

``` {r mixed modeling, error=TRUE}
#full model with all rs within breed. Right now (23 December 2024), we don't have enough variation within breeds for all these parameters, so we remove the parameters for the correlations between random slopes and intercept. 

full.choice.int.train <- glmer(choice_letter ~ condition*training + z.age + sex + first_baited_box+
             (1+ sex.c + condition.c*training.c + first_baited_box.c + z.age || breed_recoded),
             data = model.data, family = binomial,
             control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5))
)

#had to exclude correlation parameters between rss and icpt because they were unidentifiable
```


#### model output
  + Coefficients
```{r}
round(summary(full.choice.int.train)$coefficients, 3)
```
  + Individual predictors: likelihood ratio tests  
Drop1: P values for the individual effects were based on likelihood ratio tests comparing the full with respective reduced models (Barr et al., 2013; R function drop1 with argument 'test' set to "Chisq"). 

```{r}
drop1.full=drop1(full.choice.int.train, test="Chisq")
round(drop1.full,3)
#the interaction condition*breed_type is not significant, so we remove it but keep both main effects
```
#### Interaction not significant, we drop it
```{r}
full.choice.train <- glmer(choice_letter ~ condition + training + z.age + sex + first_baited_box+
             (1+ sex.c + condition.c + training.c + first_baited_box.c + z.age || breed_recoded),
             data = model.data, family = binomial,
             control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5))
)

```
#### Distribution of BLUPs
```{r}
ranef.diagn.plot(full.choice.train)
```
#### Variance Inflation Factors
```{r}
vif(full.choice.train) #collinearity is no problem
```
#### model output
  + Coefficients
```{r}
round(summary(full.choice.train)$coefficients, 3)
```
  + Individual predictors: likelihood ratio tests  
Drop1: P values for the individual effects were based on likelihood ratio tests comparing the full with respective reduced models (Barr et al., 2013; R function drop1 with argument 'test' set to "Chisq"). 

```{r}
drop1.full=drop1(full.choice.train, test="Chisq")
round(drop1.full,3)
#the interaction condition*breed_type is not significant, so we remove it but keep both main effects
```

#### confidence intervals

```{r}
boot.res.train=boot.glmm.pred(model.res=full.choice.train, excl.warnings=T,
	nboots=1000, para=T)
res.train<-round(boot.res.train$ci.estimates, 3)
res.train
```
#### model stability

```{r eval=FALSE, include=FALSE}

# One breed at a time excluded to assess the impact of possible outliers 
m.stab.train <- glmm.model.stab(model.res = full.choice.train, use = c("breed_recoded"))

table(m.stab.train$detailed$lme4.warnings) #several models failed to converge

xx.train <- as.data.frame(round(m.stab.b$summary[, -1], 3))
#table with model stability
xx.train

#plot of stability (remember to execute all three lines at the same time)
png("plots/bin_glmm03_train_stability.png")
m.stab.plot(round(m.stab.train$summary[, -1], 3))
dev.off()
```


## Inter-coder reliability for binary variable
```{r}
#LL:remove
# #define random videos to be coded (25% of final sample size = 60 videos)
# set.seed(43) #for reproducibility
# dog_ids<-sample(c(1:240), 60) #dog ids to code
# dog_ids
# reliab.data<-data.frame()
# reliab.data<-data.frame(dog_ID=dog_ids)
# 
# #create excel file for scoring reliability 
# reliab.data <- merge(reliab.data, xdata[, c("dog_ID", "dog_cdl_name")], by = "dog_ID", all.x = TRUE)
# # Restore the original order of dog_id
# reliab.data <- reliab.data[match(dog_ids, reliab.data$dog_ID), ]
# # Save reliab.data to an Excel file
# write_xlsx(reliab.data, "data/inter_rater_reliability.xlsx")

#read in reliability scorings
reliab.data<- read_excel("data/inter_rater_reliability.xlsx")

#Calculate Percentage Agreement
percentage_agreement <- sum(reliab.data$choice1 == reliab.data$choice2) / nrow(reliab.data) * 100
cat("Percentage Agreement:", round(percentage_agreement, 2), "%\n")
#Percentage Agreement: 96.67 %

#calculate Cohen's kappa
kappa_result <- kappa2(reliab.data[, c("choice1", "choice2")], "unweighted")  #Subjects = 60 Raters = 2 Kappa = 0.933 z = 7.25 p-value = 4.3e-13 
```

